---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: yes
    toc: yes
---
##Setting Global Options
```{r setoptions, echo=FALSE}
        library(knitr)
        opts_chunk$set(echo=TRUE)
        opts_chunk$set(fig.path = "/figure")
         library(lattice)
         library(ggplot2)
```


## Loading and preprocessing the data

```{r dataload}
       data<-read.csv(unz("activity.zip","activity.csv"), header=T,sep=",")
        head(data)
        data$date<-as.Date(data$date)
        
        
       
```

## What is mean total number of steps taken per day?
```{r mean total steps per day, results='asis'} 
        #aggregate number of steps per day
        ag_data<-setNames(aggregate(data$steps, by=list(data$date), sum, na.omit=T),c("Date","Steps"))
       
        #A histogram of total number of steps taken each day
        hist(ag_data$Steps, main="Total Number of Steps Taken each Day", xlab="Total Number of Steps", 
             col="royalblue")
        
        #A histogram of total number of steps per each day
        ggplot(na.omit(ag_data),aes(Date, Steps))+ geom_histogram(stat = "identity")  + 
                ggtitle("Total Number of Steps Taken per Day" )
        
         #mean total number of steps
       meansteps<- mean(ag_data[,2], na.rm=T)
        #median number of steps
       mediansteps<- median(ag_data[,2], na.rm=T)
        
```
The mean total number of steps taken per day is **`r sprintf("%.0f",meansteps)`**
The median total number of steps taken per day is **`r sprintf("%.0f",mediansteps)`**

## What is the average daily activity pattern?
```{r average daily activity pattern}
        #compute average steps per interval
        pattern<- tapply(data$steps,data$interval, mean,na.rm=T)
        
        #Make timeseries Plot
        plot(row.names(pattern), pattern, type = "l", xlab = "5-minute interval", 
             ylab = "Average Steps across all Days", 
              main = "Average Daily Activity Pattern", col = "mediumblue")
        
        #Which 5-minute interval, contains the maximum number of steps?
        maxsteps<-names(which(pattern ==max(pattern)))

```
The **`r maxsteps`**, 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

## Imputing missing values
```{r Imputing missing values}
        #Calculate the total number of missing values in the dataset (i.e. the total number of rows with NAs)
        Numnadata <- nrow(data[!complete.cases(data),])
        
        #crate a copy of the activity dataset
        filldata<-data
        
        #get indices of na data
        nadata<-which(is.na(filldata))
        
        #fill na vslues with mean number steps for the corresponding interval
        for(i in 1 : Numnadata)
        {
                index<-nadata[i]
                filldata$steps[index]<-pattern[as.character(filldata$interval[index])]
        }
          #aggregate number of steps per day of filled data
        ag_filldata<-setNames(aggregate(filldata$steps, by=list(filldata$date), sum, na.omit=T),
                              c("Date","Steps"))
       
        #A histogram of total number of steps taken each day
        hist(ag_filldata$Steps, main="Total Number of Steps Taken each Day with Filled NAs", 
             xlab="Total Number of Steps",  col="Skyblue")
        
         #mean total number of steps withfilled na
       meansteps_na<- mean(ag_filldata[,2], na.rm=T)
        #median number of steps
       mediansteps_na<- median(ag_filldata[,2], na.rm=T)
        
```
- The total number of missing values in the dataset is **`r Numnadata` **
- The mean total number of steps taken per day imputted na values is **`r sprintf("%.0f",meansteps_na)`**
- The median total number of steps taken per day with imputted na values is **`r sprintf("%.0f",mediansteps_na)`**

## Are there differences in activity patterns between weekdays and weekends?


```{r weekdays vs weekend}
        day<- weekdays(filldata$date)
        filldata$daylevel<-ifelse(day == "Saturday" | day == "Sunday", "Weekend", "Weekday")
        head(filldata)
        fillpattern<- aggregate(filldata$steps, by = list(filldata$interval, filldata$daylevel), 
        mean)
        names(fillpattern) <- c("interval", "daylevel", "steps")
        xyplot(steps ~ interval | daylevel, fillpattern, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of Steps")

```
